---
title: "DataCheckDashboard_RedLake_Flux"
author: "Victoria Martinez"
date: '2023-03-17'
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r setup, include=FALSE}
# # load flux results data from Smartflux system
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(data.table)
library(plyr)
library(cowplot)

# get windrose function from github
source(paste0("https://raw.githubusercontent.com/MargueriteM/R_functions/master/plot.windrose.R"))

# working directory (don't need to set, all file reads have full path)
# setwd("~/Desktop/OneDrive - University of Texas at El Paso/CZ_Drylands/JER_RedLakePlaya/Data/SmartFlux/results/2021/09")


# get data from summaries folder
flux.files2 <- list.files(path="C:/Users/vmartinez62/OneDrive - University of Texas at El Paso/Tower Data/JER_Playa/Data/Data_DL_Collect/SmartFlux/summaries",full.names=TRUE)

#flux.files2 <- list.files(path="C:/Users/vmartinez62/OneDrive - University of Texas at El Paso/CZO_data/Bahada/CR3000/L1/Flux/Bahada_CR3000_flux_L1_2025",full.names=TRUE)

# read the column number for each summary file
read_column_number <- function(colname){
  ret <- ncol(fread(colname, sep="\t", dec=".", header=TRUE, skip=0)[1,])
  obj_name <- tools::file_path_sans_ext(basename(colname))
  out <- data.frame(file=obj_name, colnumber=ret)
  out
}

## split into 2 because reading all at once is too big
#data1 <- ldply(flux.files2[1:100], read_column_number)

#data2 <- ldply(flux.files2[151:167], read_column_number)

## issue with files between 167-174 (zero KB files)
#data3 <- ldply(flux.files2[174:314], read_column_number)

#data4 <- ldply(flux.files2[314:500], read_column_number)

#data5 <- ldply(flux.files2[500:727], read_column_number)

#data6 <- ldply(flux.files2[727:980], read_column_number)

data7 <- ldply(flux.files2[980:1128], read_column_number)

data8 <- ldply(flux.files2[1128:1212], read_column_number)

## read the flux files as csv and combine into single dataframe
#data <- rbind(data1, data2, data3, data4, data5, data6, data7)
 
 data <- rbind(data7, data8)

# general column number is 211, select files
# 10 Oct 2024 updated smartflux to collect all variables which increased columns from 211 to 215, add 215 to filter
flux.files.read <- data %>%
  filter(colnumber==211|colnumber==215) %>%
  mutate(file.path = paste("C:/Users/vmartinez62/OneDrive - University of Texas at El Paso/Tower Data/JER_Playa/Data/Data_DL_Collect/SmartFlux/summaries/",
                          file,".txt",sep=''))

# get column names and units from complete summary files
flux.units2 <- fread(flux.files.read[flux.files.read$colnumber==211,]$file.path[1], sep="\t", dec=".", header=TRUE, skip=0, fill = TRUE)[1,]

# get data from complete summary files
#flux.data2 <- do.call("rbind", lapply(flux.files.read$file.path[1:906], header = FALSE, fread, sep="\t", dec=".",skip = 2, fill=TRUE, na.strings="NaN", col.names=colnames(flux.units2)))

flux.data2 <- do.call("rbind", lapply(flux.files.read[flux.files.read$colnumber==211,]$file.path, header = FALSE, fread, sep="\t", dec=".",skip = 2, fill=TRUE, na.strings="NaN", col.names=colnames(flux.units2)))

# format date_time variable
flux.data2 <- flux.data2 %>%
  mutate(date_time=ymd_hms(paste(date,time,sep=" ")))

# 10 Oct 2024 updated smartflux to collect all variables which increased columns from 211 to 215 read all files with 215 columns seperately
# get column names and units from complete summary files
flux.units2.av <- fread( flux.files.read[flux.files.read$colnumber==215,]$file.path[2], sep="\t", dec=".", header=TRUE, skip=0, fill = TRUE)[1,]

# get data from complete summary files
#flux.data2.av <- do.call("rbind", lapply(flux.files.read$file.path[908:nrow(flux.files.read)], header = FALSE, fread, sep="\t", dec=".",skip = 2, fill=TRUE, na.strings="NaN", col.names=colnames(flux.units2.av)))
# count total number of files with 215 columns of data,
nrow.215 <- nrow(flux.files.read[flux.files.read$colnumber==215,])
flux.data2.av <- do.call("rbind", lapply(flux.files.read[flux.files.read$colnumber==215,]$file.path[2:nrow.215], header = FALSE, fread, sep="\t", dec=".",skip = 2, fill=TRUE, na.strings="NaN", col.names=colnames(flux.units2.av)))

# format date_time variable
flux.data2.av <- flux.data2.av %>%
  mutate(date_time=ymd_hms(paste(date,time,sep=" ")))

# combine 211 column flux file with 215 row flux file
flux.data2 <- rbind(flux.data2, flux.data2.av, fill=TRUE)

# put flux.data2 in long format

flux.long <- flux.data2 %>%
  select(date, time, co2_flux, LE, H, co2_mole_fraction, h2o_mole_fraction, co2_signal_strength_7500_mean, P_RAIN_1_1_1, PPFD_1_1_1, LWOUT_1_1_1, LWIN_1_1_1, RG_1_1_1, RH_1_1_1, RN_1_1_1, SHF_1_1_1, SHF_2_1_1, SHF_3_1_1, SWC_1_1_1, SWC_1_2_1, SWC_1_3_1, SWC_1_4_1, SWC_1_5_1, SWIN_1_1_1, SWOUT_1_1_1, TA_1_1_1, TC_1_1_1, TS_1_1_1, TS_1_2_1, TS_1_3_1, TS_1_4_1, TS_1_5_1, TS_2_1_1, TS_3_1_1, VIN_1_1_1, WD_1_1_1, ALB_1_1_1, air_pressure) %>%
  pivot_longer(!c(date, time), names_to="variable",values_to="value") %>%
   mutate(date_time=ymd_hms(paste(date,time,sep=" ")))


# Create data for footprint graphing   
footprint.data <- flux.data2 %>%
  select(date_time,WD_1_1_1,`u*`,`x_90%`,`x_70%`,`x_50%`,`x_30%`,`x_10%`)%>%
  filter(`u*`>0.2) %>%
  drop_na
          
```
# Plot any variables

Column {.sidebar}
-----------------------------

### Select Date Range and Variables

These graphs show most recent incoming SmartFlux data from the tower at Red Lake Playa. Data are 30min mean values.

**Note: Flux calculations are preliminary and for visualisation only!!** 

#### Last timestamp data updated:
```{r, echo=FALSE, warning=FALSE, message=FALSE}

print(max(flux.long$date_time))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}

dateRangeInput("daterange", "Date range:",
                 start = min(flux.long$date),
                 end   = max(flux.long$date),
                 min = min(flux.long$date),
                 max = max(flux.long$date))

selectInput("variable1", "Flux variable 1:",
             c("CO2 Flux" = "co2_flux",
               "Latent Heat Exhange" = "LE",
               "Sensible Heat Flux" = "H",
               "CO2 Mole Fraction" = "co2_mole_fraction",
               "H2O Mole Fraction" = "h2o_mole_fraction",
               "CO2 Signal Strength" = "co2_signal_strength_7500_mean",
               "Rain" = "P_RAIN_1_1_1",
               "Photosynthetic Photon Flux Density" = "PPFD_1_1_1",
               "Long Wave Radiation Out" = "LWOUT_1_1_1",
               "Long Wave Radiation In" = "LWIN_1_1_1",
               "Short Wave Radiation Out" = "SWOUT_1_1_1",
               "Short Wave Radiation In" = "SWIN_1_1_1"))

selectInput("variable2", "Flux variable 2:",
             c("CO2 Flux" = "co2_flux",
               "Latent Heat Exhange" = "LE",
               "Sensible Heat Flux" = "H",
               "CO2 Signal Strength" = "co2_signal_strength_7500_mean",
               "Rain" = "P_RAIN_1_1_1",
               "Photosynthetic Photon Flux Density" = "PPFD_1_1_1",
               "Long Wave Radiation Out" = "LWOUT_1_1_1",
               "Long Wave Radiation In" = "LWIN_1_1_1",
               "Short Wave Radiation Out" = "SWOUT_1_1_1",
               "Short Wave Radiation In" = "SWIN_1_1_1"))

selectInput("variable3", "Flux variable 3:",
             c("CO2 Flux" = "co2_flux",
               "Latent Heat Exhange" = "LE",
               "Sensible Heat Flux" = "H",
               "CO2 Signal Strength" = "co2_signal_strength_7500_mean",
               "Rain" = "P_RAIN_1_1_1",
               "Photosynthetic Photon Flux Density" = "PPFD_1_1_1",
               "Long Wave Radiation Out" = "LWOUT_1_1_1",
               "Long Wave Radiation In" = "LWIN_1_1_1",
               "Short Wave Radiation Out" = "SWOUT_1_1_1",
               "Short Wave Radiation In" = "SWIN_1_1_1"))
```

Column
-------------------------
Graph three variables in parallel
```{r, echo = FALSE, warning = FALSE, message = FALSE}
renderPlot({
  c <- flux.long %>%
    filter(date >= input$daterange[[1]] & date <= input$daterange[[2]] &
           variable==input$variable1)
  
  ggplot(c, aes(date_time, value)) + 
    labs(y=input$variable1)+
     geom_line()+
    geom_point(size=0.7, color="red")+
    theme_bw()})
     

renderPlot({
  c <- flux.long %>%
    filter(date >= input$daterange[[1]] & date <= input$daterange[[2]] &
           variable==input$variable2)
  
  ggplot(c, aes(date_time, value)) +
    labs(y=input$variable2)+
    geom_line()+
    geom_point(size=0.7, color="red")+
    theme_bw()})

renderPlot({
  c <- flux.long %>%
    filter(date >= input$daterange[[1]] & date <= input$daterange[[2]] &
           variable==input$variable3)
  
  ggplot(c, aes(date_time, value)) +
    labs(y=input$variable3)+
    geom_line()+
    geom_point(size=0.7, color="red")+
    theme_bw()})
```

# Plot Wind Data

Column {.sidebar}
-----------------------------
  
  ### Select Date Range and Variables
  
  These graphs show most recent incoming SmartFlux data from the tower at Red Lake Playa. Data are 30min mean values.

**Note: Flux calculations are preliminary and for visualisation only!!** 
  
  
```{r, echo=FALSE, warning=FALSE, message=FALSE}

dateRangeInput("daterange2", "Date range:",
               start = min(flux.long$date),
               end   = max(flux.long$date),
               min = min(flux.long$date),
               max = max(flux.long$date))
```

Column
-------------------------
  
```{r, echo = FALSE, warning = FALSE, message = FALSE}
renderPlot({wind.dat <- flux.data2%>%
  select(date,WS_1_1_1,WD_1_1_1, wind_speed, wind_dir)%>%
  drop_na()%>%
  filter((date >= input$daterange2[[1]] & date <= input$daterange2[[2]]))

wind.2d <- plot.windrose(wind.dat,
                         wind.dat$WS_1_1_1,
                         wind.dat$WD_1_1_1)+
  theme_bw()+
  labs(title="2-D Anemometer")

#
wind.3d <- plot.windrose(wind.dat,
                         wind.dat$wind_speed,
                         wind.dat$wind_dir)+
  theme_bw()+
  labs(title="Sonic Anemometer")

# graph 2-D and Sonic Wind Rose side-by-side
plot_grid(wind.2d,wind.3d,labels=c("A","B"))
})
```

# Plot Footprint Data

Column
-------------------------

```{r, echo=FALSE, warning=FALSE, message=FALSE}

dateRangeInput("daterange2", "Date range:",
               start = min(flux.long$date),
               end   = max(flux.long$date),
               min = min(flux.long$date),
               max = max(flux.long$date)) 

# use windrose to plot footprints
renderPlot({
  
  # use windrose function to make a footprint graph: ... it's a hack.
  p1 <- plot.windrose(footprint.data,footprint.data$`x_90%`,footprint.data$WD_1_1_1,spdmax=1000,spdres=100)+
    theme_bw()+
    labs(title="Footprint direction & distance")
  
  p1})

# histogram of footprint distance
renderPlot({ggplot(footprint.data)+
    geom_histogram(aes(`x_10%`),fill="blue")+
    geom_histogram(aes(`x_30%`),fill="red")+
    geom_histogram(aes(`x_50%`),fill="green")+
    geom_histogram(aes(`x_70%`),fill="purple")+
    geom_histogram(aes(`x_90%`),colour="grey")+
    labs(x="Distance Contribution (m)")+
    theme_bw()})
```

# Focus on Soil Data

Column {.sidebar}
---------------------------------------

### Select Date Range and variables

These graphs show most recent incoming Biomet data from the tower at Red Lake Playa. Data are 30min mean values.

**Note: Flux calculations are preliminary and for visualisation only!!** 

#### Last timestamp data updated:

```{r, echo=FALSE, warning=FALSE, message=FALSE}

print(max(flux.long$date_time))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}

dateRangeInput("daterange3", "Date range:",
                 start = min(flux.long$date),
                 end   = max(flux.long$date),
                 min = min(flux.long$date),
                 max = max(flux.long$date))

selectInput("variable4", "Biomet variable:",
             c("Atmospheric Pressure" = "^air_pressure",
               "Air Temperature" = "^TA",
               "Canopy Temperature" ="^TC",
               "Rainfall" =  "^P_RAIN",
               "Longwave Radiation" = "^LWIN|^LWOUT",
               "Shortwave Radiation" = "^SWIN|^SWOUT",
               "Albedo" = "^ALB",
              "Global Radiation & PAR" = "^RG|^PPFD",
              "Net Radiation" = "^RN"))

selectInput("variable5", "Soil variable 1:",
             c("All Soil Temperature Profile" = "^TS_1",
               "All Soil Temperature Integrated (0.5-15cm)" = "^TS_2|^TS_3",
               "All Soil Moisture Profile" = "^SWC",
               "All Soil Heat Flux (16cm)" ="^SHF",
               "Soil Temperature 10cm" ="^TS_1_1_1",
               "Soil Temperature 20cm" ="^TS_1_2_1",
               "Soil Temperature 31cm" ="^TS_1_3_1",
               "Soil Temperature 49cm" ="^TS_1_4_1",
               "Soil Temperature 93cm" ="^TS_1_5_1",
               "Soil Moisture 10cm" ="^SWC_1_1_1",
               "Soil Moisture 20cm" ="^SWC_1_2_1",
               "Soil Moisture 31cm" ="^SWC_1_3_1",
               "Soil Moisture 49cm" ="^SWC_1_4_1",
               "Soil Moisture 93cm" ="^SWC_1_5_1"))

selectInput("variable6", "Soil Variable 2:",
            c("All Soil Temperature Profile" = "^TS_1",
               "All Soil Temperature Integrated (0.5-15cm)" = "^TS_2|^TS_3",
               "All Soil Moisture Profile" = "^SWC",
               "All Soil Heat Flux (16cm)" ="^SHF",
              "Soil Temperature 10cm" ="^TS_1_1_1",
               "Soil Temperature 20cm" ="^TS_1_2_1",
               "Soil Temperature 31cm" ="^TS_1_3_1",
               "Soil Temperature 49cm" ="^TS_1_4_1",
               "Soil Temperature 93cm" ="^TS_1_5_1",
              "Soil Moisture 10cm" ="^SWC_1_1_1",
               "Soil Moisture 20cm" ="^SWC_1_2_1",
               "Soil Moisture 31cm" ="^SWC_1_3_1",
               "Soil Moisture 49cm" ="^SWC_1_4_1",
               "Soil Moisture 93cm" ="^SWC_1_5_1"))

radioButtons("ax.scales", "Axis scales:",
             c("Fixed" = "fixed", 
               "Free x" = "free_x",
               "Free y" = "free_y",
               "Free x & y" = "free"))
```

Column
-------------------------------------------------------------
Graph Met and Soil Variables in Parallel  
Profile measurements are at 10, 20, 31, 49, 93cm
```{r, echo = FALSE, warning = FALSE, message = FALSE}

# biomet plot
renderPlot({
    c <-  flux.long %>%
  filter(str_detect(variable,input$variable4) &
           (date >= input$daterange3[[1]] & date <= input$daterange3[[2]])) 
  
   # d <- plot.titles %>%
   #   filter(selection1 == input$variable4)
   #  # setnames(c,input$variable1,"selected")
   # 
    # plot
     ggplot(c, aes(date_time, value))+
    geom_line()+
    #labs(title = d$name, y=d$name)+
    facet_grid(variable~., scales = input$ax.scales)+
 theme_bw()})

# soil plot 1
renderPlot({
    c <-  flux.long %>%
  filter(str_detect(variable,input$variable5) &
           (date >= input$daterange3[[1]] & date <= input$daterange3[[2]])) 
  
   # d <- plot.titles %>%
   #   filter(selection1 == input$variable5)
   #  # setnames(c,input$variable1,"selected")
 
    # plot
     ggplot(c, aes(date_time, value, colour=variable))+
    geom_line()+
    #labs(title = d$name, y=d$name)+
    facet_grid(variable~., scales = input$ax.scales)+
 theme_bw()+
       theme(legend.position="bottom")})

# soil plot 2
renderPlot({
    c <-  flux.long %>%
  filter(str_detect(variable,input$variable6) &
           (date >= input$daterange3[[1]] & date <= input$daterange3[[2]])) 
  
   # d <- plot.titles %>%
   #   filter(selection1 == input$variable5)
   #  # setnames(c,input$variable1,"selected")
 
    # plot
     ggplot(c, aes(date_time, value, colour=variable))+
    geom_line()+
    #labs(title = d$name, y=d$name)+
       facet_grid(input$variable3~.)+
 theme_bw()+
           facet_grid(variable~., scales = input$ax.scales)+
       theme(legend.position = "bottom")})

```